
% ex.tex - Colin Ramsay - 22 Oct 99
%
% The wrapper for the appendix of examples.
%
%   5   10   15   20   25   30   35   40   45   50   55   60   65   70   75
% ..|....|....|....|....|....|....|....|....|....|....|....|....|....|....|

In this appendix we give some examples of \ace\ runs.
A stand-alone discussion of some of the features of these runs is included,
  although parts of these runs are mentioned in the body of the text, as 
  illustrations of specific features of \ace's behaviour.
The \ttt{ex***.in} files supplied as part of this documentation can be
  used to run these examples, although an example may be presented as if it
  were generated interactively, and the output may be edited for reasons of
  space or perspicuity.
There may be minor variations in the exact format of the output, since
  \ace\ is continually being `improved'\kern-1pt.
Unless otherwise noted, all parameters are defaulted and the default
  build of \ace\ was used.
In multipart runs, note that parameters from an earlier part may carry
  across to a later one.
Note that some of the examples may require a machine with a large amount
  of memory.

\section{Getting started}\label{ex000}
\input{ex000.tex}

\section{Emulating Sims}\label{ex001}
\input{ex001.tex}

\section{Row filling}\label{ex002}
\input{ex002.tex}

\section{Equivalent presentations}

TBA: $F(2,7)$, using \ttt{rep} \amp \ttt{aep} ...

\section{Deduction queues}

TBA: ... (see test009)

\section{Large enumerations}

Suppose that the presentation given is such that the final coset table
  exceeds the 4\,Gbyte limit imposed by 32-bit machines; e.g., an index of
  $250 \times 10^6$\kern-2pt, with a 5-column table and 4 byte integers.
We are justified in regarding such an enumeration as `big'\kern-1.5pt, 
  since it will require more than 4\,Gbyte of storage no matter how
  efficiently it is performed.
Of course, even trivial enumerations may exceed this limit if they are
  very pathological (see, e.g., \cite{HR2}).
However, we have no (easy) way of knowing whether or not such enumerations
  can be done within the 4\,Gbyte limit, so we are hesitant to classify
  them as big.
\ace\ is 64-bit `aware'\kern-1.5pt, and can use more than 4\,Gbyte of
  memory if it is available.
Note however that the number of cosets (i.e., the number of rows in the
  coset table) is still limited by the size of a signed \ttt{int}.
So the maximum size of a table is $2^{31} - n$ cosets, where $n$ is
  probably $3$; one since we can't actually represent $+2147483648$, one
  since coset \#0 is not used, and one since we need to count one past the
  top of the table.

Some trivial group enumerations involving more than 1\,G total cosets 
  and 4\,Gbyte of memory were reported in \cite{HR2}.
However, the first big enumeration, in the above sense, done by \ace\ was
  the Thomson simple group $\mathit{Th}$ \cite{MR1829480};
$|\mathit{Th}|=2^{15}.3^{10}.5^3.7^2.13.19.31$,
and it contains $^3D_4(2):3$ as an index $143.127.000$ subgroup.
TBA: ...

\section{Looping}

TBA: ...

\section{Use of \ttt{st}}\label{ex007}

TBA: ...

\section{Use of \ttt{cy}, \ttt{nc}, \ttt{cc} and \ttt{rc}} 

TBA: ...

