%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%W  moreexamples.tex    ACE appendix - more examples    Joachim Neub"user
%W                                                            Greg Gamble
%%
%H  $Id$
%%
%Y  Copyright (C) 2000  Centre for Discrete Mathematics and Computing
%Y                      Department of Computer Science & Electrical Eng.
%Y                      University of Queensland, Australia.
%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Chapter{More Examples}

The examples in this chapter  are  intended  to  provide  the  nearest
{\GAP} equivalent of the similarly named  sections  in  Appendix~A  of
`ace3001.dvi' (the standalone manual in  directory  `standalone-doc').
It is *strongly* recommended that users  unfamiliar  with  the  {\ACE}
standalone avoid this chapter until they have gained some  familiarity
with the {\ACE} standalone. There is a *lot* of detail here, which the
novice {\ACE} Share Package user won't want  to  know  about.  Please,
despite the name of the  first  section  of  this  chapter,  read  the
examples in Appendix~"Examples" first.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Getting Started}

Each     of     the      functions      `ACECosetTableFromGensAndRels'
(see~"ACECosetTableFromGensAndRels"), `ACEStats'  (see~"ACEStats"  ---
non-interactive  version)  and  `ACEStart'  (see~"ACEStart"),  may  be
called with three arguments: <fgens> (the  group  generators),  <rels>
(the group relators), and <sgens> (the subgroup generators). While  it
is legal for the arguments <rels> and <sgens> to be empty lists, it is
always an error for <fgens> to be empty, e.g.

\beginexample
gap> ACEStats([],[],[]);
Error, fgens (arg[1]) must be a non-empty list of group generators ...
 called from
CALL_ACE( "ACEStats", arg[1], arg[2], arg[3] ) called from
<function>( <arguments> ) called from read-eval-loop
Entering break read-eval-print loop ...
 type: 'quit;' to quit to outer loop, or
 type: 'fgens := <val>; return;' to assign <val> to fgens to continue.
brk> fgens := FreeGeneratorsOfFpGroup(FreeGroup("a"));
[ a ]
brk> return;
rec( index := 0, cputime := 45, cputimeUnits := "10^-2 seconds", 
  activecosets := 499998, maxcosets := 499998, totcosets := 499998 )
gap>
\endexample

The example shows that the {\ACE} share package does allow you to recover
from the `break'-loop, but let us start again.

After defining generators, we can do an enumeration:

\beginexample
gap> F := FreeGroup("a", "b");;
gap> fgens := FreeGeneratorsOfFpGroup(F);
[ a, b ]
\endexample

\atindex{ACEStats}{@\noexpand`ACEStats'!example}
*An `ACEStats' example*

By  default,  the  presentation  is  not  echoed;   use   the   `echo'
(see~"option echo") option if you want that.  Also,  by  default,  the
{\ACE} binary only prints a *results message*, but we won't  see  that
unless   we   set   `InfoACE'   to   a   level   of   at    least    2
(see~"SetInfoACELevel"):

\beginexample
gap> SetInfoACELevel(2);
\endexample

Calling `ACEStats' with arguments `fgens', `[]', `[]', defines a  free
froup with 2 generators, since the second argument  defines  an  empty
relator list; and since  the  third  argument  is  an  empty  list  of
generators, the  subgroup  defined  is  trivial.  So  the  enumeration
overflows:

\beginexample
gap> ACEStats(fgens, [], []);
#I  OVERFLOW (a=249998 r=83333 h=83333 n=249999; l=337 c=0.29; m=249998 t=2499\
98)
rec( index := 0, cputime := 29, cputimeUnits := "10^-2 seconds", 
  activecosets := 249998, maxcosets := 249998, totcosets := 249998 )
\endexample

The  line  starting  with   ```\#I  '''.  is  the  `Info'-ed  *results
message* from {\ACE};  see  Appendix~"The  Meanings  of  ACE's  Output
Messages" for details  on  what  it  means.  Observe  that  since  the
enumeration overflowed, {\ACE}'s result message  has  been  translated
into a {\GAP} record with `index' field 0.

To dump  out  the  presentation  and  parameters  associated  with  an
enumeration,  {\ACE}  provides  the  `sr'  (see~"option  sr")  option.
However, you won't see output of this  command,  unless  you  set  the
`InfoACELevel' to at least 3. Also, to ensure the reliability  of  the
output of the `sr' option, an enumeration  should  *precede*  it;  for
`ACEStats' (and `ACECosetTableFromGensAndRels') the directive  `start'
(see~"option start") required to initiate an enumeration  is  inserted
(automatically) after all the  user's  options,  except  if  the  user
herself supplies an option that initiates an enumeration (namely,  one
of `start' or `begin' (see~"option start"), `aep'  (see~"option  aep")
or `rep' (see~"option rep")). Interactively,  the  equivalent  of  the
`sr' command is `ACEParameters' (see~"ACEParameters"), which gives  an
output record that is immediately {\GAP}-usable.  With  the  above  in
mind let's  rerun  the  enumeration  and  get  {\ACE}'s  dump  of  the
presentation and parameters:

\beginexample
gap> SetInfoACELevel(3);
gap> ACEStats(fgens, [], [] : start, sr := 1);
#I  ACE 3.000        Sat Mar  3 21:05:53 2001
#I  =========================================
#I  Host information:
#I    name = hamal
#I  OVERFLOW (a=249998 r=83333 h=83333 n=249999; l=337 c=0.29; m=249998 t=2499\
98)
#I    #-- ACE 3.000: Run Parameters ---
#I  Group Name: G;
#I  Group Generators: ab;
#I  Group Relators: ;
#I  Subgroup Name: H;
#I  Subgroup Generators: ;
#I  Wo:1000000; Max:249998; Mess:0; Ti:-1; Ho:-1; Loop:0;
#I  As:0; Path:0; Row:1; Mend:0; No:0; Look:0; Com:10;
#I  C:0; R:0; Fi:7; PMod:3; PSiz:256; DMod:4; DSiz:1000;
#I    #--------------------------------
#I  ***
rec( index := 0, cputime := 29, cputimeUnits := "10^-2 seconds", 
  activecosets := 249998, maxcosets := 249998, totcosets := 249998 )
\endexample

Observe that at `InfoACE' level 3, one also gets {\ACE}'s  banner.  We
could have printed out the first few lines of the coset  table  if  we
had wished, using the `print' (see~"option print") option, but note as
with the `sr' option, an enumeration should *precede* it. Here's  what
happens if you disregard this (recall, we  still  have  the  `InfoACE'
level set to 3):

\beginexample
#I  ACE 3.000        Sat Mar  3 21:06:42 2001
#I  =========================================
#I  Host information:
#I    name = hamal
#I  ** ERROR (continuing with next line)
#I     no information in table
#I  OVERFLOW (a=249998 r=83333 h=83333 n=249999; l=337 c=0.29; m=249998 t=2499\
98)
#I  ***
rec( index := 0, cputime := 29, cputimeUnits := "10^-2 seconds", 
  activecosets := 249998, maxcosets := 249998, totcosets := 249998 )
\endexample

Essentially, because {\ACE} had  not  done  an  enumeration  prior  to
getting the `print' directive, it complained with an  ```**  ERROR''',
recovered  and  went  on  with  the  `start'  directive  automatically
inserted by the `ACEStats' command:  no  ill  effects  at  the  {\GAP}
level, but also no table.

Now, let's do what we should have done (to get those first  few  lines
of the coset table), namely, insert  the  `start'  option  before  the
`print' option (the `InfoACE' level is still 3):

\beginexample
gap> ACEStats(fgens, [], [] : start, print := [-1, 12]);
#I  ACE 3.000        Sat Mar  3 21:07:50 2001
#I  =========================================
#I  Host information:
#I    name = hamal
#I  OVERFLOW (a=249998 r=83333 h=83333 n=249999; l=337 c=0.29; m=249998 t=2499\
98)
#I  co: a=249998 r=83333 h=83333 n=249999; c=+0.00
#I   coset ||      a      A      b      B   order   rep've
#I  -------+---------------------------------------------
#I       1 ||      2      3      4      5
#I       2 ||      6      1      7      8       0   a
#I       3 ||      1      9     10     11       0   A
#I       4 ||     12     13     14      1       0   b
#I       5 ||     15     16      1     17       0   B
#I       6 ||     18      2     19     20       0   aa
#I       7 ||     21     22     23      2       0   ab
#I       8 ||     24     25      2     26       0   aB
#I       9 ||      3     27     28     29       0   AA
#I      10 ||     30     31     32      3       0   Ab
#I      11 ||     33     34      3     35       0   AB
#I      12 ||     36      4     37     38       0   ba
#I  ***
rec( index := 0, cputime := 29, cputimeUnits := "10^-2 seconds", 
  activecosets := 249998, maxcosets := 249998, totcosets := 249998 )
\endexample

The values we gave to the `print' option, told  {\ACE}  to  print  the
first 12 lines and include coset  representatives.  Note  that,  since
there are no relators, the table has separate  columns  for  generator
inverses. So the default workspace of $1000000$ words allows  a  table
of $249998 = 1000000/4 - 2$ cosets. Since row  `fill'ing  (see~"option
fill") is on by default, the table is simply  filled  with  cosets  in
order. Note that a compaction phase is done before printing the table,
but that this does nothing here (the lowercase `co' tag), since  there
are no dead cosets. The coset representatives are simply all  possible
freely reduced words, in length plus lexicographic (i.e. `lenlex'; see
Section~"Coset Table Standardisation Schemes") order.

\atindex{ACECosetTableFromGensAndRels}%
{@\noexpand`ACECosetTableFromGensAndRels'!example}
*Using `ACECosetTableFromGensAndRels'*

The  essential  difference  between  the  functions   `ACEStats'   and
`ACECosetTableFromGensAndRels' is that `ACEStats' parses the  *results
message* from the {\ACE} binary and outputs a {\GAP} record containing
statistics  of  the  enumeration,  and  `ACECosetTableFromGensAndRels'
after parsing the *results message*, goes on to parse  {\ACE}'s  coset
table, if it can, and outputs a {\GAP} list of lists version  of  that
table. So, if we had used  `ACECosetTableFromGensAndRels'  instead  of
`ACEStats' in our examples  above,  we  would  have  observed  similar
output, except that we would have ended up in a `break'-loop  (because
the enumeration overflows) instead of obtaining  a  record  containing
enumeration statistics. We have already seen an  example  of  that  in
Section~"Using ACE Directly to Generate a Coset Table".  So,  here  we
will consider two options that prevent one  entering  a  `break'-loop,
namely   the   `silent'   (see~"option   silent")   and   `incomplete'
(see~"option  incomplete")  options.  Firstly,  let's  take  the  last
`ACEStats' example, but use `ACECosetTableFromGensAndRels' instead and
include the `silent' option. (We still have the `InfoACE' level set at
3.)

\beginexample
gap> ACECosetTableFromGensAndRels(fgens, [], [] : start, print := [-1, 12],
>                                                 silent);
#I  ACE 3.000        Sat Mar  3 21:09:10 2001
#I  =========================================
#I  Host information:
#I    name = hamal
#I  OVERFLOW (a=249998 r=83333 h=83333 n=249999; l=337 c=0.29; m=249998 t=2499\
98)
#I  co: a=249998 r=83333 h=83333 n=249999; c=+0.00
#I   coset ||      a      A      b      B   order   rep've
#I  -------+---------------------------------------------
#I       1 ||      2      3      4      5
#I       2 ||      6      1      7      8       0   a
#I       3 ||      1      9     10     11       0   A
#I       4 ||     12     13     14      1       0   b
#I       5 ||     15     16      1     17       0   B
#I       6 ||     18      2     19     20       0   aa
#I       7 ||     21     22     23      2       0   ab
#I       8 ||     24     25      2     26       0   aB
#I       9 ||      3     27     28     29       0   AA
#I      10 ||     30     31     32      3       0   Ab
#I      11 ||     33     34      3     35       0   AB
#I      12 ||     36      4     37     38       0   ba
#I  ***
fail
\endexample

Since, the enumeration overflowed and the  `silent'  option  was  set,
`ACECosetTableFromGensAndRels' simply returned `fail'.  But  hang  on,
{\ACE} at least has a partial table; we should be able to get that  in
{\GAP},  in  a  situation  like  this.  We  can.  We  simply  use  the
`incomplete' option, instead of the `silent' option.  However,  if  we
did that with the example above, the result would be an enormous table
(the number of *active cosets* is 249998); so  let  us  also  set  the
`max' (see~"option max") option, in order that we should  get  a  more
modestly sized partial table. Finally, we  will  use  `print  :=  -12'
since it is a shorter equivalent alternative to `print :=  [-1,  12]'.
Note that the output here was obtained with {\GAP}  4.3.  With  {\GAP}
4.2 the output is similar  except  that  the  last  `Info'-ed  message
(before the final output)  states  that  the  coset  table  result  is
incomplete only, since no standardisation is done. It turns  out  that
the table displayed via the `print'  option  is  already  in  `lenlex'
standard form; so despite the differences in the {\GAP} versions, both
{\GAP} 4.2 and {\GAP} 4.3 output the same table.

\beginexample
gap> ACECosetTableFromGensAndRels(fgens, [], [] : max := 12,
>                                                 start, print := -12,
>                                                 incomplete);
#I  ACE 3.000        Sat Mar  3 21:10:28 2001
#I  =========================================
#I  Host information:
#I    name = hamal
#I  OVERFLOW (a=12 r=4 h=4 n=13; l=5 c=0.00; m=12 t=12)
#I  co: a=12 r=4 h=4 n=13; c=+0.00
#I   coset ||      a      A      b      B   order   rep've
#I  -------+---------------------------------------------
#I       1 ||      2      3      4      5
#I       2 ||      6      1      7      8       0   a
#I       3 ||      1      9     10     11       0   A
#I       4 ||     12      0      0      1       0   b
#I       5 ||      0      0      1      0       0   B
#I       6 ||      0      2      0      0       0   aa
#I       7 ||      0      0      0      2       0   ab
#I       8 ||      0      0      2      0       0   aB
#I       9 ||      3      0      0      0       0   AA
#I      10 ||      0      0      0      3       0   Ab
#I      11 ||      0      0      3      0       0   AB
#I      12 ||      0      4      0      0       0   ba
#I  ***
#I  co: a=12 r=4 h=4 n=13; c=+0.00
#I   coset ||      a      A      b      B
#I  -------+----------------------------
#I       1 ||      2      3      4      5
#I       2 ||      6      1      7      8
#I       3 ||      1      9     10     11
#I       4 ||     12      0      0      1
#I       5 ||      0      0      1      0
#I       6 ||      0      2      0      0
#I       7 ||      0      0      0      2
#I       8 ||      0      0      2      0
#I       9 ||      3      0      0      0
#I      10 ||      0      0      0      3
#I      11 ||      0      0      3      0
#I      12 ||      0      4      0      0
#I  ACECosetTable: Coset table is incomplete, reduced and lenlex standardised.
[ [ 2, 6, 1, 12, 0, 0, 0, 0, 3, 0, 0, 0 ], 
  [ 3, 1, 9, 0, 0, 2, 0, 0, 0, 0, 0, 4 ], 
  [ 4, 7, 10, 0, 1, 0, 0, 2, 0, 0, 3, 0 ], 
  [ 5, 8, 11, 1, 0, 0, 2, 0, 0, 3, 0, 0 ] ]
\endexample

Observe, that despite the fact that {\ACE} is  able  to  define  coset
representatives for all 12 coset numbers  defined,  the  body  of  the
coset table now contains a 0 at each  place  formerly  occupied  by  a
coset number larger than 12 (0 essentially represents ``don't know'').
To get a table that is the same in the first  12  rows  as  before  we
would have had to set `max' to 38, since that was  the  largest  coset
number that appeared in the body of  the  12-line  table,  previously.
Also, note that the `max' option *preceded* the `start' option;  since
the interface respects the order in which options are put by the user,
the enumeration invoked by `start'  would  otherwise  have  only  been
restricted by the size of `workspace'  (see~"option  workspace").  The
warning that the coset table is incomplete is emitted at `InfoACE'  or
`InfoWarning' level 1, i.e.~by default, you will see it.

\atindex{ACEStart}{@\noexpand`ACEStart'!example}
*Using `ACEStart'*

The     limitation     of     the     functions     `ACEStats'     and
`ACECosetTableFromGensAndRels' (on three arguments) is  that  they  do
not interact with {\ACE}; they call {\ACE}  with  user-defined  input,
and collect and parse the output for  either  statistics  or  a  coset
table. On the other hand,  the  `ACEStart'  (see~"ACEStart")  function
allows one to start up an {\ACE} process and maintain a dialogue  with
it. Moreover, via the functions `ACEStats' and `ACECosetTable'  (on  1
or no arguments), one is able to extract the same information that  we
could with the non-interactive versions of these  functions.  However,
we can also do a lot more. Each {\ACE}  option  that  provides  output
that can be used from within {\GAP} has  a  corresponding  interactive
interface function that parses and translates that output into a  form
usable from within {\GAP}.

Now we emulate our  (successful)  `ACEStats'  exchanges  above,  using
interactive  {\ACE}  interface  functions.  We  could  do  this  with:
`ACEStart(0, fgens, [], [] : start, sr := 1);'  where  the  `0'  first
argument tells `ACEStart' not to  insert  `start'  after  the  options
explicitly listed. Alternatively, we may do the following  (note  that
the `InfoACE' level is still 3):

\beginexample
#I  ACE 3.000        Sat Mar  3 21:12:28 2001
#I  =========================================
#I  Host information:
#I    name = hamal
#I  ***
#I  OVERFLOW (a=249998 r=83333 h=83333 n=249999; l=337 c=0.30; m=249998 t=2499\
98)
1
gap> ACEParameters(1);
#I    #-- ACE 3.000: Run Parameters ---
#I  Group Name: G;
#I  Group Generators: ab;
#I  Group Relators: ;
#I  Subgroup Name: H;
#I  Subgroup Generators: ;
#I  Wo:1000000; Max:249998; Mess:0; Ti:-1; Ho:-1; Loop:0;
#I  As:0; Path:0; Row:1; Mend:0; No:0; Look:0; Com:10;
#I  C:0; R:0; Fi:7; PMod:3; PSiz:256; DMod:4; DSiz:1000;
#I    #--------------------------------
rec( enumeration := "G", subgroup := "H", workspace := 1000000, 
  max := 249998, messages := 0, time := -1, hole := -1, loop := 0, asis := 0, 
  path := 0, row := 1, mendelsohn := 0, no := 0, lookahead := 0, 
  compaction := 10, ct := 0, rt := 0, fill := 7, pmode := 3, psize := 256, 
  dmode := 4, dsize := 1000 )
\endexample

Observe that the `ACEStart' call returned an integer (1, here). All  8
forms of the `ACEStart' function, return an  integer  that  identifies
the interactive {\ACE} interface function  initiated  or  communicated
with. We may use this integer to tell any interactive {\ACE} interface
function which interactive {\ACE} process we wish to communicate with.
Above we passed `1' to the `ACEParameters' command which caused `sr :=
1' (see~"option sr") to be passed to the  interactive  {\ACE}  process
indexed by 1 (the process we just started), and  a  record  containing
the  parameter  options  (see  Section~"ACE  Parameter  Options")   is
returned. Note that the ``Run Parameters'': `Group Generators', `Group
Relators'  and   `Subgroup   Generators'   are   considered   ``args''
(i.e.~arguments) and a record containing  these  is  returned  by  the
`GetACEArgs' (see~"GetACEArgs")  command;  or  they  may  be  obtained
individually     via      the      commands:      `ACEGroupGenerators'
(see~"ACEGroupGenerators"),  `ACERelators'   (see~"ACERelators"),   or
`ACESubgroupGenerators' (see~"ACESubgroupGenerators").

We can obtain the enumeration statistics record, via  the  interactive
version of `ACEStats' (see~"ACEStats!interactive") :

\beginexample
gap> ACEStats(1); # The interactive version of ACEStats takes 1 or no arg'ts
rec( index := 0, cputime := 30, cputimeUnits := "10^-2 seconds", 
  activecosets := 249998, maxcosets := 249998, totcosets := 249998 )
\endexample

To display 12 lines of the  coset  table  with  coset  representatives
without invoking a further enumeration we could do: `ACEStart(0,  1  :
print   :=   [-1,   12]);'.   Alternatively,   we    may    use    the
`ACEDisplayCosetTable' (see~"ACEDisplayCosetTable") (the table  itself
is emitted at `InfoACE' level 1, since by default we  presumably  want
to see it):

\beginexample
gap> ACEDisplayCosetTable(1, [-1, 12]);
#I  co: a=249998 r=83333 h=83333 n=249999; c=+0.00
#I   coset ||      a      A      b      B   order   rep've
#I  -------+---------------------------------------------
#I       1 ||      2      3      4      5
#I       2 ||      6      1      7      8       0   a
#I       3 ||      1      9     10     11       0   A
#I       4 ||     12     13     14      1       0   b
#I       5 ||     15     16      1     17       0   B
#I       6 ||     18      2     19     20       0   aa
#I       7 ||     21     22     23      2       0   ab
#I       8 ||     24     25      2     26       0   aB
#I       9 ||      3     27     28     29       0   AA
#I      10 ||     30     31     32      3       0   Ab
#I      11 ||     33     34      3     35       0   AB
#I      12 ||     36      4     37     38       0   ba
#I  ------------------------------------------------------------
\endexample

Still with the same interactive {\ACE} process we can now emulate  the
`ACECosetTableFromGensAndRels' exchange that  gave  us  an  incomplete
coset table. Note that it is still necessary to invoke an  enumeration
after setting the `max' (see~"option max") option. We could just  call
`ACECosetTable' with the argument 1 and the same 4 options we used for
`ACECosetTableFromGensAndRels'.   Alternatively,   we   can   do   the
equivalent of the 4  options  one  (or  two)  at  a  time,  via  there
equivalent interactive commands.  Note  that  the  `ACEStart'  command
(without `0' as first argument) inserts a `start' directive after  the
user option `max':

\beginexample
gap> ACEStart(1 : max := 12);
#I  ***
#I  OVERFLOW (a=12 r=4 h=4 n=13; l=5 c=0.00; m=12 t=12)
1
\endexample

Now the following `ACEDisplayCosetTable' command does  the  equivalent
of the `print := [-1, 12]' option.

\beginexample
gap> ACEDisplayCosetTable(1, [-1, 12]);
#I  co: a=12 r=4 h=4 n=13; c=+0.00
#I   coset ||      a      A      b      B   order   rep've
#I  -------+---------------------------------------------
#I       1 ||      2      3      4      5
#I       2 ||      6      1      7      8       0   a
#I       3 ||      1      9     10     11       0   A
#I       4 ||     12      0      0      1       0   b
#I       5 ||      0      0      1      0       0   B
#I       6 ||      0      2      0      0       0   aa
#I       7 ||      0      0      0      2       0   ab
#I       8 ||      0      0      2      0       0   aB
#I       9 ||      3      0      0      0       0   AA
#I      10 ||      0      0      0      3       0   Ab
#I      11 ||      0      0      3      0       0   AB
#I      12 ||      0      4      0      0       0   ba
#I  ------------------------------------------------------------
\endexample

Finally, we call `ACECosetTable' with 1 argument  (which  invokes  the
interactive version of `ACECosetTableFromGensAndRels') with the option
`incomplete'.

\beginexample
gap> ACECosetTable(1 : incomplete);          
#I  start = yes, continue = yes, redo = yes
#I  ***
#I  OVERFLOW (a=12 r=4 h=4 n=13; l=4 c=0.00; m=12 t=12)
#I  co: a=12 r=4 h=4 n=13; c=+0.00
#I   coset |      a      A      b      B
#I  -------+----------------------------
#I       1 |      2      3      4      5
#I       2 |      6      1      7      8
#I       3 |      1      9     10     11
#I       4 |     12      0      0      1
#I       5 |      0      0      1      0
#I       6 |      0      2      0      0
#I       7 |      0      0      0      2
#I       8 |      0      0      2      0
#I       9 |      3      0      0      0
#I      10 |      0      0      0      3
#I      11 |      0      0      3      0
#I      12 |      0      4      0      0
#I  ACECosetTable: Coset table is incomplete, reduced & lenlex standardised.
[ [ 2, 6, 1, 12, 0, 0, 0, 0, 3, 0, 0, 0 ], 
  [ 3, 1, 9, 0, 0, 2, 0, 0, 0, 0, 0, 4 ], 
  [ 4, 7, 10, 0, 1, 0, 0, 2, 0, 0, 3, 0 ], 
  [ 5, 8, 11, 1, 0, 0, 2, 0, 0, 3, 0, 0 ] ]
\endexample

Observe the line beginning ```\#I start = yes,'''  (the  first line in
the output of `ACECosetTable'). This line appears in response  to  the
option `mode' (see~"option mode") inserted  by  `ACECosetTable'  after
any user options; it is inserted  in  order  to  check  that  no  user
options  (possibly  made  before  the   `ACECosetTable'   call)   have
invalidated {\ACE}'s coset table. Since the line also says `continue =
yes', the mode `continue' (the least expensive  of  the  three  modes;
see~"option continue") is directed at {\ACE} which evokes  a  *results
message*. Then `ACECosetTable' extracts the  incomplete  table  via  a
`print' (see "option print") directive. If you wish  to  see  all  the
options that are directed to {\ACE}, set  the  `InfoACE'  level  to  4
(then all such commands are `Info'-ed behind a  ```ToACE> '''  prompt;
see~"SetInfoACELevel").

We now set things up to do the alternating group $A_5$, of order $60$.
We turn  messaging  on  via  the  `messages'  (see~"option  messages")
option, but set the interval high enough so that we will only get  the
``Run Parameters'' block (the same as that obtained with  `sr  :=  1')
and no progress messages. Recall `F' is the free group we  defined  on
generators `fgens': `"a"' and `"b"', and the `InfoACE' level is  still
3. Note, that when {\GAP} prints  `F.1'  ($=$ `fgens[1]') it  displays
`a', but the *variable* `a' is (at  the  moment)  unassigned;  so  for
convenience (in defining relators, for example) we  first  assign  the
variable `a' to be `F.1' (and `b' to be `F.2').

\beginexample
gap> a := F.1;; b := F.2;;
gap> # Enumerating A_5 = < a, b || a^2, b^3, (a*b)^5 > 
gap> # over Id (trivial subgp)
gap> ACEStart(1, fgens, [a^2, b^3, (a*b)^5], [] 
>                                     # 4th arg empty (to define Id)
>             : enumeration := "A_5", # Define the Group Name
>               subgroup := "Id",     # Define the Subgroup Name
>               max := 0,             # Set `max' back to default (no limit)
>               messages := 1000);    # Progress messages every 1000 iter'ns
#I  ***
#I  ***
#I    #-- ACE 3.000: Run Parameters ---
#I  Group Name: A_5;
#I  Group Generators: ab;
#I  Group Relators: (a)^2, (b)^3, (ab)^5;
#I  Subgroup Name: Id;
#I  Subgroup Generators: ;
#I  Wo:1000000; Max:333331; Mess:1000; Ti:-1; Ho:-1; Loop:0;
#I  As:0; Path:0; Row:1; Mend:0; No:3; Look:0; Com:10;
#I  C:0; R:0; Fi:6; PMod:3; PSiz:256; DMod:4; DSiz:1000;
#I    #--------------------------------
#I  INDEX = 60 (a=60 r=77 h=1 n=77; l=3 c=0.00; m=66 t=76)
1
\endexample

Since we  did  not  specify  a  strategy  the  `default'  (see~"option
default") strategy was followed and  hence  coset  number  definitions
were R (i.e.~HLT) style, and a total of $76$  coset  numbers  (`t=76')
were defined. Note, that {\ACE} already ``knew'' the group  generators
and subgroup generators, using the `relators' (see~"option  relators")
option:

\beginexample
gap> ACEStart(1 : relators := ToACEWords(fgens, [a^2, b^3, (a*b)^5]),
>                 enumeration := "A_5",                                        
>                 subgroup := "Id",                                            
>                 max := 0,                                                    
>                 messages := 1000);                                           
#I  Detected usage of a synonym of one (or more) of the options:
#I      `group', `relators', `generators'.
#I  Discarding current values of args.
#I  (The new args will be extracted from ACE, later).
#I  No group generators saved. Setting value(s) from ACE ...
#I    #-- ACE 3.000: Run Parameters ---
#I  Group Name: A_5;
#I  Group Generators: ab;
#I  Group Relators: (a)^2, bbb, ababababab;
#I  Subgroup Name: Id;
#I  Subgroup Generators: ;
#I  Wo:1000000; Max:333331; Mess:1000; Ti:-1; Ho:-1; Loop:0;
#I  As:0; Path:0; Row:1; Mend:0; No:3; Look:0; Com:10;
#I  C:0; R:0; Fi:6; PMod:3; PSiz:256; DMod:4; DSiz:1000;
#I    #--------------------------------
#I  ***
#I    #-- ACE 3.000: Run Parameters ---
#I  Group Name: A_5;
#I  Group Generators: ab;
#I  Group Relators: (a)^2, (b)^3, (ab)^5;
#I  Subgroup Name: Id;
#I  Subgroup Generators: ;
#I  Wo:1000000; Max:333331; Mess:1000; Ti:-1; Ho:-1; Loop:0;
#I  As:0; Path:0; Row:1; Mend:0; No:3; Look:0; Com:10;
#I  C:0; R:0; Fi:6; PMod:3; PSiz:256; DMod:4; DSiz:1000;
#I    #--------------------------------
#I  INDEX = 60 (a=60 r=77 h=1 n=77; l=3 c=0.00; m=66 t=76)
1
\endexample

Note the usage  of  `ToACEWords'  (see~"ToACEWords")  to  provide  the
appropriate string value of the `relators' option. Also,  observe  the
`Info'-ed warning of the action  triggered  by  using  the  `relators'
option, that says that the current values of the  ``args''  (i.e.~what
would be returned by `GetACEArgs'; see~"GetACEArgs")  were  discarded,
which immediately triggered the action of reinstantiating the value of
`ACEData.io[1].args' (which is what the `Info':

\begintt
#I  No group generators saved. Setting value(s) from ACE ...
\endtt

was all about). Also observe that the  ``Run  Parameters''  block  was
`Info'-ed twice; the first time was due to  `ACEStart'  emitting  `sr'
with value `1' to {\ACE} ({\ACE}'s response is used to  re-instantiate
`ACEData.io[1].args').

In  particular,  {\GAP}  no  longer  thinks  `fgens'  are  the   group
generators:

\beginexample
gap> ACEGroupGenerators(1)=fgens;
false
\endexample

Groan! We will just have to re-instantiate everything:

\beginexample
gap> fgens := ACEGroupGenerators(1);;                      
gap> F := GroupWithGenerators(fgens);; a := F.1;; b := F.2;;
\endexample

We will now define a non-trivial subgroup, and monitor all the actions
of the enumerator, by setting `messages := 1'.  As  for  defining  the
relators, we could use the 1-argument version of `ACEStart', in  which
case we would use the `subgroup' (see~"option subgroup")  option  with
the value `ToACEWords(fgens, [ a*b ])'. Instead, we will use the  more
convenient 4-argument form:

\beginexample
gap> ACEStart(1, ACEGroupGenerators(1), ACERelators(1), [ a*b ]
>             : messages := 1);
#I  ***
#I  ***
#I    #-- ACE 3.000: Run Parameters ---
#I  Group Name: A_5;
#I  Group Generators: ab;
#I  Group Relators: (a)^2, (b)^3, (ab)^5;
#I  Subgroup Name: Id;
#I  Subgroup Generators: ab;
#I  Wo:1000000; Max:333331; Mess:1; Ti:-1; Ho:-1; Loop:0;
#I  As:0; Path:0; Row:1; Mend:0; No:3; Look:0; Com:10;
#I  C:0; R:0; Fi:6; PMod:3; PSiz:256; DMod:4; DSiz:1000;
#I    #--------------------------------
#I  AD: a=2 r=1 h=1 n=3; l=1 c=+0.00; m=2 t=2
#I  SG: a=2 r=1 h=1 n=3; l=1 c=+0.00; m=2 t=2
#I  RD: a=3 r=1 h=1 n=4; l=2 c=+0.00; m=3 t=3
#I  RD: a=4 r=2 h=1 n=5; l=2 c=+0.00; m=4 t=4
#I  RD: a=5 r=2 h=1 n=6; l=2 c=+0.00; m=5 t=5
#I  RD: a=6 r=2 h=1 n=7; l=2 c=+0.00; m=6 t=6
#I  RD: a=7 r=2 h=1 n=8; l=2 c=+0.00; m=7 t=7
#I  RD: a=8 r=2 h=1 n=9; l=2 c=+0.00; m=8 t=8
#I  RD: a=9 r=2 h=1 n=10; l=2 c=+0.00; m=9 t=9
#I  CC: a=8 r=2 h=1 n=10; l=2 c=+0.00; d=0
#I  RD: a=9 r=5 h=1 n=11; l=2 c=+0.00; m=9 t=10
#I  RD: a=10 r=5 h=1 n=12; l=2 c=+0.00; m=10 t=11
#I  RD: a=11 r=5 h=1 n=13; l=2 c=+0.00; m=11 t=12
#I  RD: a=12 r=5 h=1 n=14; l=2 c=+0.00; m=12 t=13
#I  RD: a=13 r=5 h=1 n=15; l=2 c=+0.00; m=13 t=14
#I  RD: a=14 r=5 h=1 n=16; l=2 c=+0.00; m=14 t=15
#I  CC: a=13 r=6 h=1 n=16; l=2 c=+0.00; d=0
#I  CC: a=12 r=6 h=1 n=16; l=2 c=+0.00; d=0
#I  INDEX = 12 (a=12 r=16 h=1 n=16; l=3 c=0.00; m=14 t=15)
1
\endexample

Observe that we used `ACERelators(1)' (see~"ACERelators") to grab  the
value  of  the  relators  we  had  defined  earlier.  We   also   used
`ACEGroupGenerators(1)' (see~"ACEGroupGenerators") to  get  the  group
generators (`fgens' would do instead, except if, as  in  our  sequence
above, the `relators' option was used, since  this  causes  {\GAP}  to
lose  the  knowledge  that  `fgens'  and  `ACEGroupGenerators(1)'  are
equal).

The  run  ended  with  12  active   (see   Section~"Coset   Statistics
Terminology") coset numbers (`a=12') after defining a total number  of
15 coset numbers (`t=15'). So there must  have  been  3  coincidences:
observe  that  there  were  3  progress  messages  (see  Appendix~"The
Meanings of ACE's Output Messages") with a `CC:' tag.

We now dump out the  statistics  accumulated  during  the  run,  using
`ACEDumpStatistics'  (see~"ACEDumpStatistics"),  which   `Info's   the
{\ACE}  output  of  the  `statistics'  (see~"option  statistics")   at
`InfoACE' level 1.

\beginexample
#I    #- ACE 3.000: Level 0 Statistics --
#I  cdcoinc=0 rdcoinc=2 apcoinc=0 rlcoinc=0 clcoinc=0
#I    xcoinc=2 xcols12=4 qcoinc=3
#I    xsave12=0 s12dup=0 s12new=0
#I    xcrep=6 crepred=0 crepwrk=0 xcomp=0 compwrk=0
#I  xsaved=0 sdmax=0 sdoflow=0
#I  xapply=1 apdedn=1 apdefn=1
#I  rldedn=0 cldedn=0
#I  xrdefn=1 rddedn=5 rddefn=13 rdfill=0
#I  xcdefn=0 cddproc=0 cdddedn=0 cddedn=0
#I    cdgap=0 cdidefn=0 cdidedn=0 cdpdl=0 cdpof=0
#I    cdpdead=0 cdpdefn=0 cddefn=0
#I    #----------------------------------
\endexample

The statistic `qcoinc=3' states what we had already observed,  namely,
that there  were  three  coincidences.  Of  these,  two  were  primary
coincidences  (`rdcoinc=2').  Since  `t=15',   there   were   fourteen
non-trivial coset number definitions; one was during  the  application
of coset 1 to the subgroup generator (`apdefn=1'), and  the  remainder
occurred during applications of the  coset  numbers  to  the  relators
(`rddefn=13'). For more details on the meanings of the  variables  you
will need to read the C code comments.

%Note how the pre-printout compaction phase now  does  some  work  (the
%upper-case `CO' tag), since there were coincidences,  and  hence  dead
%cosets. Note how `b/B' have been used as the first two columns,  since
%these must be occupied by  a  generator/inverse  pair  or  a  pair  of
%involutions. The `a' column is also the  `A'  column,  as  `a'  is  an
%involution.
%
%\begintt
%  print  TABLE : -1, 12 ;
%CO: a=12 r=13 h=1 n=13; c=+0.00
% coset |      b      B      a   order   rep've
%-------+--------------------------------------
%     1 |      3      2      2
%     2 |      1      3      1       3   B
%     3 |      2      1      4       3   b
%     4 |      8      5      3       5   ba
%     5 |      4      8      6       2   baB
%     6 |      9      7      5       5   baBa
%     7 |      6      9      8       3   baBaB
%     8 |      5      4      7       5   bab
%     9 |      7      6     10       5   baBab
%    10 |     12     11      9       3   baBaba
%    11 |     10     12     12       2   baBabaB
%    12 |     11     10     11       3   baBabab
%\endtt
%
%If we define the generator order to be that of the columns,  then  the
%table above is not in canonic form, and the coset representatives  are
%not in order. We now standardise the table; note the compaction  phase
%before standardisation, although it does nothing  in  this  particular
%case. Note how, if we read through the table in row-major  order,  new
%cosets are introduced using the smallest available  number,  and  that
%the representatives are now in order and are minimal.
%
%\begintt
%st;
%co/ST: a=12 r=13 h=1 n=13; c=+0.00
%pr:-1,12;
%co: a=12 r=13 h=1 n=13; c=+0.00
% coset |      b      B      a   order   rep've
%-------+--------------------------------------
%     1 |      2      3      3
%     2 |      3      1      4       3   b
%     3 |      1      2      1       3   B
%     4 |      5      6      2       5   ba
%     5 |      6      4      7       5   bab
%     6 |      4      5      8       2   baB
%     7 |      8      9      5       5   baba
%     8 |      9      7      6       5   baBa
%     9 |      7      8     10       3   babaB
%    10 |     11     12      9       3   babaBa
%    11 |     12     10     12       3   babaBab
%    12 |     10     11     11       2   babaBaB
%\endtt
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%E
