%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%W  otheroptions.tex    ACE appendix - other options     Alexander Hulpke
%W                                                      Joachim Neub"user
%W                                                            Greg Gamble
%%
%H  $Id$
%%
%Y  Copyright (C) 2000  Centre for Discrete Mathematics and Computing
%Y                      Department of Computer Science & Electrical Eng.
%Y                      University of Queensland, Australia.
%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Chapter{Other ACE Options}

Here we will list  all  the  known  {\ACE}  options  that  users  will
normally only wish to use if generating an input file,  by  using  the
option `aceinfile' (see~"option aceinfile"). Most of the options below
are  for  interactive  use  with  the  standalone;  from  the   {\GAP}
interface, we provide other functions for interactive use.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Experimentation Options}

\beginitems

\>`aep:=<val>'{option aep}@{option `aep'}&
Runs  the enumeration for `a'll `e'quivalent `p'resentations;
<val> is in the integer range 0 to 7.

The `aep' option runs  an  enumeration  for  combinations  of  relator
ordering, relator rotations, and relator inversions.

The argument <val> is considered as a binary number.  Its  three  bits
are treated as flags, and control relator rotations (the  $2^0$  bit),
relator inversions (the $2^1$ bit) and relator  orderings  (the  $2^2$
bit),  respectively;  where  $1$  means  \lq{}active'  and  $0$  means
\lq{}inactive'. (See below for an example).

The `aep' option first performs a \lq{}priming run' using the  options
as they stand. In particular, the `asis' and  `messages'  options  are
honoured.

It then turns `asis' on and `messages' off  (i.e.~sets  `messages'  to
0), and generates and tests the  requested  equivalent  presentations.
The maximum and minimum values attained by `m' (the maximum number  of
coset numbers defined at any stage) and `t' (the total number of coset
numbers defined) are tracked, and each  time  a  new  \lq{}record'  is
found, the relators used and the summary result line is  printed.  See
Appendix~"The Meanings of ACE's output messages" for a  discussion  of
the statistics `m' and `t'. To observe these messages either  set  the
`InfoLevel' of `InfoACE' to 3 or non-interactively you can peruse  the
{\ACE} output file (see~"option aceoutfile").

The order in which the  equivalent  presentations  are  generated  and
tested has no particular significance, but note that the  presentation
as given *after* the initial priming run) is the  *last*  presentation
to be generated and tested, so that  the  group's  relators  are  left
`unchanged' by running the `aep' option, (not that  a  non-interactive
user cares).

As discussed by Cannon, Dimino, Havas  and  Watson  \cite{CDHW73}  and
Havas and Ramsay \cite{HR99b} such equivalent presentations can  yield
large variations in  the  number  of  coset  numbers  required  in  an
enumeration. For this command, we are interested in this variation.

After  the  final  presentation  is  run,   some   additional   status
information messages are printed to the {\ACE} output file:

\beginlist
\item{--}  the number of runs which yielded a finite index; 
\item{--}  the total number of runs (excluding the priming run); and 
\item{--}  the range of values observed for `m' and `t'.
\endlist

As an example (drawn from the discussion in \cite{HR99a}) consider the
enumeration   of   the   $448$   coset   numbers   of   the   subgroup
$\langle  a^2,Ab \rangle$ of the group
$$ (8,7 \mid 2,3) 
    = \langle a,b \mid a^8 = b^7 = (ab)^2 = (Ab)^3 = 1 \rangle. $$
There are $4!=24$  relator  orderings  and  $2^4=16$  combinations  of
relator or inverted relator. Exponents are  taken  into  account  when
rotating relators, so the relators given give rise to 1, 1,  2  and  2
rotations respectively, for a total of $1.1.2.2=4$  combinations.  So,
for  `aep'  =  $7$   (resp.~$3$),   $24.16.4=1536$   (resp.~$16.4=64$)
equivalent presentations are tested.

*Notes:*
There is no way to stop the `aep'  option  before  it  has  completed,
other than killing the task. So do a reality check beforehand  on  the
size of the search space and the time for each enumeration. If you are
interested in  finding  a  \lq{}good'  enumeration,  it  can  be  very
helpful, in terms of running time, to put a tight limit on the  number
of coset numbers via the `max'  option.  You  may  also  have  to  set
`compaction' = $100$ to prevent time-wasting attempts to recover space
via compaction. This maximises throughput  by  causing  the  \lq{}bad'
enumerations, which are in  the  majority,  to  overflow  quickly  and
abort. If you wish to explore  a  very  large  search-space,  consider
firing up many copies of {\ACE}, and starting each with a \lq{}random'
equivalent  presentation.  Alternatively,  you  could  use  the  `rep'
command.

Interactively,             use              `ACEAllEquivPresentations'
(see~"ACEAllEquivPresentations").

\>`rep:=<val>'{option rep}@{option `rep'}
\>`rep:=[<val>, <Npresentations>]'{option rep}@{option `rep'}&
Runs  the enumeration for `r'andom `e'quivalent `p'resentations;
<val> is in the integer range 0 to 7;
<Npresentations> must be a positive integer.

The `rep' (random equivalent  presentations)  option  complements  the
`aep'  option.  It  generates  and  tests   some   random   equivalent
presentations. The argument <val>  acts  as  for  `aep'.  It  is  also
possible to set the number <Npresentations>  of  random  presentations
used (by default, eight  are  used),  by  using  the  extended  syntax
`rep:=[<val>,<Npresentations>]'.

The routine first  turns  `asis'  on  and  `messages'  off  (i.e.~sets
`messages' to 0), and then generates and tests the requested number of
random equivalent presentations. For each presentation,  the  relators
used and the  summary  result  line  are  printed.  To  observe  these
messages either set the `InfoLevel' of `InfoACE'  to  at  least  3  or
non-interactively you can peruse the {\ACE} output  file  (see~"option
aceoutfile").

*Notes:*
The relator inversions and rotations are \lq{}genuinely'  random.  The
relator permuting is a little bit of a kludge, with the  \lq{}quality'
of the permutations tending to improve with successive  presentations.
When the `rep' command  completes,  the  presentation  active  is  the
*last* one generated, (not that the non-interactive user cares).

*Guru Notes:*
It might appear that neglecting to restore the  original  presentation
is an error. In fact, it is a useful feature! Suppose that  the  space
of equivalent presentations is too  large  to  exhaustively  test.  As
noted in the entry for `aep', we can start up multiple copies of `aep'
at random points in the  search-space.  Manually  generating  `random'
equivalent presentations to serve as starting-points  is  tedious  and
error-prone. The `rep' option provides a simple solution;  simply  run
`rep := 7' before `aep := 7'.

Interactively,            use            `ACERandomEquivPresentations'
(see~"ACERandomEquivPresentations").

\enditems

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Options that Modify a Presentation}

\beginitems

\>`generators:=<wordList>'{option generators}@{option `generators'}&
Define the subgroup `generators'; <wordList> must be a list  of  words
in the group generators.
(Shortest abbreviation: `gen'.)

There should *never* be a need to use this command, even when creating
a standalone input file. The subgroup generators should  be  input  as
one of the arguments of an  {\ACE}  interface  function.  By  default,
there are no subgroup generators and the  subgroup  is  trivial.  This
command allows a list of subgroup generating words to be entered.

\>`group:=<list>'{option group}@{option `group'}&
Define the `group' generators; <list> must be a list of  {\GAP}  group
generators.
(Shortest abbreviation: `gr'.)

There should *never* be a need to use this command, even when creating
a standalone input file. The group generators should be input  as  one
of the arguments of an {\ACE} interface function.  If  the  generators
each have names that as strings are single  lowercase  letters,  those
same strings are used to represent  the  same  generators  by  {\ACE};
otherwise,  {\ACE}  will  represent  each  generator  by  an  integer,
numbered sequentially from 1.

*Notes:*
Any use of the  `group'  command  which  actually  defines  generators
invalidates any previous enumeration, and stays in  effect  until  the
next `group' command. Any words for the  group  or  subgroup  must  be
entered using the nominated generator format, and  all  printout  will
use this format. A valid set of generators is the minimum  information
necessary before {\ACE} will attempt an enumeration.

*Guru Notes:*
The columns of the coset table are allocated in the same order as  the
generators are listed, insofar as this is  possible,  given  that  the
first two columns must be  a  generator/inverse  pair  or  a  pair  of
involutions. The ordering of the columns can, in  some  cases,  affect
the definition sequence of cosets and  impact  the  statistics  of  an
enumeration.

\>`relators:=<wordList>'{option relators}@{option `relators'}&
Define the group `relators'; <wordList> must be a list of words in the
group generators.
(Shortest abbreviation: `rel'.)

There should *never* be a need to use this command, even when creating
a standalone input file. The group relators should be input as one  of
the arguments of an {\ACE} interface function.  If  <wordList>  is  an
empty list, the group is free.

\>`sg:=<wordList>'{option sg}@{option `sg'}&
Adds the words in <wordList> to any  `s'ubgroup  `g'enerators  already
present; <wordList> must be a list of words in the group generators.

The enumeration must  be  (re)`start'ed  or  `redo'ne,  it  cannot  be
`continue'd.

Interactively,             use              `ACEAddSubgroupGenerators'
(see~"ACEAddSubgroupGenerators").

\>`rl:=<wordList>'{option rl}@{option `rl'}&
Appends the `r'elator `l'ist,  <wordList>  to  the  existent  list  of
relators present; <wordList> must be a list  of  words  in  the  group
generators.

The enumeration must  be  (re)`start'ed  or  `redo'ne,  it  cannot  be
`continue'd.

Interactively, use `ACEAddRelators' (see~"ACEAddRelators").

\>`ds:=<list>'{option ds}@{option `ds'}&
`D'elete `s'ubgroup generators; <list> must  be  a  list  of  positive
integers.

This command  allows  subgroup  generators  to  be  deleted  from  the
presentation. If the generators are numbered from 1 in the output  of,
say, the `sr' command (see~"option sr"), then the generators listed in
`<list>' are deleted; `<list>' must be a strictly increasing sequence.

Interactively,            use            `ACEDeleteSubgroupGenerators'
(see~"ACEDeleteSubgroupGenerators").

\>`dr:=<list>'{option dr}@{option `dr'}&
`D'elete relators; <list> must be a list of positive integers.

This  command  allows  group  relators  to   be   deleted   from   the
presentation. If the relators are numbered from 1 in  the  output  of,
say, the `sr' command (see~"option sr"), then the relators  listed  in
`<list>' are deleted; `<list>' must be a strictly increasing sequence.

Interactively, use `ACEDeleteRelators' (see~"ACEDeleteRelators").

\>`cc:=<val>'{option cc}@{option `cc'}&
Make `c'oset <val> `c'oincide with coset 1; <val> should be a positive
integer.

Prints out the representative of coset `<val>', and  adds  it  to  the
subgroup generators; i.e., forces coset `<val>' to coincide with coset
1, the subgroup.

\>`rc:=<val>'{option rc}@{option `rc'}
\>`rc:=[<val>]'{option rc}@{option `rc'}
\>`rc:=[<val>, <attempts>]'{option rc}@{option `rc'}&
Enforce `r'andom `c'oincidences; <val> and <attempts> must be positive
integers.

This option attempts upto <attempts> (or,  in  the  first  and  second
forms, 8) times to find nontrivial subgroups with index a multiple  of
<val> by repeatedly making random coset numbers coincident with  coset
1 and seeing what happens. The starting coset table must be non-empty,
but need not be complete. For each attempt, we repeatedly  add  random
coset representatives to the subgroup and `redo' the  enumeration.  If
the table becomes too small, the  attempt  is  aborted,  the  original
subgroup generators restored, and another attempt made. If an  attempt
succeeds, then the new set of subgroup generators is retained.

Interactively, use `ACERandomCoincidences' (see~"ACERandomCoincidences").

*Guru Notes:*
A coset number can have many different coset representatives. Consider
running `standard' before `rc', to canonicise the table and hence  the
coset representatives.

\enditems

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Mode Options}

\beginitems

\>`mode'{option mode}@{option `mode'}&
Prints the possible enumeration `mode's.
(Shortest abbreviation: `mo'.)

Prints the possible enumeration  `mode's  (i.e.~which  of  `continue',
`redo' or `start' are possible (see~"option continue",  "option  redo"
and "option start").

\>`begin'{option begin}@{option `begin'}
\>`start'{option start}@{option `start'}&
Start an enumeration. (Shortest abbreviation of `begin' is `beg'.)

Any existing information in the table is cleared, and the  enumeration
starts from coset 1 (i.e., the subgroup).

Interactively, use `ACEStart' (see~"ACEStart").

\>`check'{option check}@{option `check'}
\>`redo'{option redo}@{option `redo'}&
`Redo' an extant enumeration, using the current parameters.

As opposed to `start' (see~"option start"), which clears  an  existing
coset table, any existing information in the table  is  retained,  and
the enumeration is restarted from coset 1 (i.e., the subgroup).

Interactively, use `ACERedo' (see~"ACERedo").

*Notes:*
This option is really intended for the case where additional  relators
(option `rl'; see~"option  rl")  and/or  subgroup  generators  (option
`sg'; see~"option sg") have been introduced. The current table,  which
may be incomplete  or  exhibit  a  finite  index,  is  still  *valid*.
However, the additional data may allow the enumeration to complete, or
cause a collapse to a smaller index.

\>`continue'{option continue}@{option `continue'}&
`Continue' the current enumeration, building upon the existing table.
(Shortest abbreviation: `cont'.)

If a previous run stopped without producing a finite index you can, in
principle, change any of the parameters and `continue' on. Of  course,
if you make any changes which invalidate the current table, you  won't
be allowed to `continue',  although  you  may  be  allowed  to  `redo'
(see~"option redo"). If `redo' is not  allowed,  you  must  re-`start'
(see~"option start").

Interactively, use `ACEContinue' (see~"ACEContinue").

\enditems

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Options that Interact with the Operating System}

\beginitems

\>`ai'{option ai}@{option `ai'}
\>`ai:=<filename>'{option ai}@{option `ai'}&
`A'lter `i'nput to standard input or <filename>; <filename> must be  a
string.

By default, commands to {\ACE} are read from standard input (i.e., the
keyboard). With no value `ai' causes {\ACE} to revert to reading  from
standard input; otherwise, the `ai' command closes the  current  input
file,  and  opens  `<filename>'  as  the  source   of   commands.   If
`<filename>' can't be opened, input reverts to standard input.

*Notes:*
If you switch to taking input from (another) file, remember to  switch
back before the end of that file; otherwise the `EOF' there will cause
{\ACE} to terminate.

\>`bye'{option bye}@{option `bye'}
\>`exit'{option exit}@{option `exit'}
\>`qui'{option qui}@{option `qui'}&
Quit {\ACE}. (Shortest abbreviation of `qui' is `q'.)

This quits {\ACE} nicely, printing the date and  the  time.  An  `EOF'
(end-of-file;  i.e.,  `\^{}d')  has  the  same   effect,   so   proper
termination occurs if {\ACE} is taking its input from a script file.

Interactively, use `ACEQuit' (see~"ACEQuit").

Note  that  `qui'  actually  abbreviates  the   corresponding   {\ACE}
directive `quit', but since `quit' is  a  {\GAP}  keyword  it  is  not
available via the {\GAP} interface to {\ACE}.

\>`system:=<string>'{option system}@{option `system'}&
Do a shell escape, to execute <string>; <string> must be a string.
(Shortest abbreviation: `sys'.)

Since {\GAP} already provides `Exec()' for this purpose,  this  option
is unlikely to have a use.

\enditems

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Query Options}

\beginitems

\>`cycles'{option cycles}@{option `cycles'}&
Print out the table in `cycles'. (Shortest abbreviation: `cy'.)

This option prints out the permutation representation.

Interactively, use `ACECycles' (see~"ACECycles").

\>`dump'{option dump}@{option `dump'}
\>`dump:=<level>'{option dump}@{option `dump'}
\>`dump:=[<level>]'{option dump}@{option `dump'}
\>`dump:=[<level>, <detail>]'{option dump}@{option `dump'}&
`Dump's the internal variables of {\ACE}; <level> must be  an  integer
in the range 0 to 2, and <detail> must be 0 or 1.
(Shortest abbreviation: `d'.)

The value of <level> determines which of the three levels of {\ACE} to
dump. (You will need to read the standalone manual to understand  what
Levels 0, 1 and 2 are all about.) The value of <detail> determines the
amount of detail (`<detail> = 0' means less detail).  The  first  form
(with no arguments) selects `<level> = 0, <detail> =  0'.  The  second
form of this command makes `<detail> = 0'. This option is intended for
gurus; the source code should be consulted  to  see  what  the  output
means. Non-interactively, the output from `dump' is  directed  to  the
temporary file with path `ACEData.banner'.

Interactively, use `ACEDumpVariables' (see~"ACEDumpVariables").

\>`help'{option help}@{option `help'}&
Print the {\ACE} help screen. (Shortest abbreviation: `h'.)

This option prints the list of options of the {\ACE} binary. Note that
this list is longer than a standard screenful.

\>`nc'{option nc}@{option `nc'}
\>`nc:=<val>'{option nc}@{option `nc'}
\>`normal'{option normal}@{option `normal'}
\>`normal:=<val>'{option normal}@{option `normal'}&
Check or attempt to enforce normal closure; <val> must be 0 or 1.

This option tests the subgroup for normal closure within the group. If
a conjugate of a subgroup generator by a generator, is  determined  to
belong to a coset other than coset 1, it is printed out, and if `<val>
=  1',  then  any  such  conjugate  is  also  added  to  the  subgroup
generators. With no argument or if `<val> = 0', {\ACE}  does  not  add
any new subgroup generators.

*Notes:*
The method of determination of  whether  a  conjugate  of  a  subgroup
generator is in the subgroup, is by testing whether it can  be  traced
from coset 1 to coset 1  (see  `trace':~"option trace").

The resultant subgroup need not be  normally  closed  after  executing
option `nc'  with  the  value  1.  It  is  still  possible  that  some
conjugates of the newly added subgroup generators will not be elements
of the subgroup.

Interactively,       use       `ACEConjugatesForSubgroupNormalClosure'
(see~"ACEConjugatesForSubgroupNormalClosure").

\>`options'{option options}@{option `options'}&
Dump version information of the {\ACE} binary.
(Shortest abbreviation: `opt'.)

A rather unfortunate name for an option; this command dumps details of
the \lq{}options' included in the version of {\ACE}  when  the  {\ACE}
binary was compiled.

Use  the   command   `ACEVersion();'   (see~"ACEVersion")   for   this
information, instead, unless you want it in an {\ACE} standalone input
file.

A typical output, is as follows:

\begintt
Executable built:
  Sat Feb 27 15:57:59 EST 1999
Level 0 options:
  statistics package = on
  coinc processing messages = on
  dedn processing messages = on
Level 1 options:
  workspace multipliers = decimal
Level 2 options:
  host info = on
\endtt

\>`oo:=<val>'{option oo}@{option `oo'}
\>`order:=<val>'{option order}@{option `order'}&
Print a coset representative of a coset number with order  a  multiple
of <val> modulo the subgroup; <val> must be an integer.

This option finds a coset with order a multiple  of  `|<val>|'  modulo
the subgroup, and prints out its coset representative.  If  `<val>  \<
0', then all coset numbers meeting the  requirement  are  printed.  If
`<val> > 0', then just the first coset number meeting the  requirement
is printed. Also, `<val> = 0' is permitted; this special value effects
the printing of the orders (modulo the subgroup) of all coset numbers.

Interactively, use `ACEOrders' (see~"ACEOrders"), for the case  `<val>
= 0', or `ACEOrder' (see~"ACEOrder"), otherwise.

\>`sr'{option sr}@{option `sr'}
\>`sr:=<val>'{option sr}@{option `sr'}&
Print out parameters of the current presentation; <val> must be  0  or
1.

No argument, or `<val> = 0', prints out the  `Group  Name',  `Subgroup
Name', the group's `relators'  and  the  subgroup's  `generators'.  If
`<val> = 1', then the current setting of the \lq{}run  parameters'  is
also printed. The printout is the same as that produced at  the  start
of a run when option `messages' (see~"option messages") is non-zero.

Interactively,  use  `ACEGroupGenerators'  (see~"ACEGroupGenerators"),
`ACERelators'   (see~"ACERelators"),   `ACESubgroupGenerators'    (see
"ACESubgroupGenerators"), and `ACEParameters' (see~"ACEParameters").

*Notes:*
Only use *after* an enumeration run; otherwise, the value 0  for  some
options will be unreliable.

\>`print'{option print}@{option `print'}
\>`print:=<val>'{option print}@{option `print'}
\>`print:=[<val>]'{option print}@{option `print'}
\>`print:=[<val>, <last>]'{option print}@{option `print'}
\>`print:=[<val>, <last>, <by>]'{option print}@{option `print'}&
Compact and print the coset table;  <val>  must  be  an  integer,  and
<last> and <by> must be positive integers.
(Shortest abbreviation: `pr'.)

In the first (no value) form, `print' prints the entire  coset  table,
without orders or coset  representatives.  In  the  second  and  third
forms, the absolute value of <val> is taken to be the last line of the
table to be printed (and 1 is taken to be the first);  in  the  fourth
and fifth forms, `|<val>|' is taken to be the first line of the  table
to be printed, and <last> is taken to be the number of the  last  line
to be printed. In the last  form,  the  table  is  printed  from  line
`|<val>|' to line <last> in steps of <by>. If <val> is negative,  then
the  orders   modulo   the   subgroup   (if   available)   and   coset
representatives are printed also.

\>`sc:=<val>'{option sc}@{option `sc'}
\>`stabilising:=<val>'{option stabilising}@{option `stabilising'}&
Print out the coset numbers whose elements stabilise  (i.e.~normalise)
the subgroup; <val> must be an integer.
(Shortest abbreviation of `stabilising' is `stabil'.)

If `<val> > 0', the first `<val>' non-trivial (i.e.~other  than  coset
1) coset numbers whose elements stabilise the subgroup are printed. If
`<val> = 0', all non-trivial coset numbers  whose  elements  stabilise
the subgroup, plus their representatives, are printed.  If  `<val>  \<
0', the first  `|<val>|'  non-trivial  coset  numbers  whose  elements
stabilise the subgroup, plus their representatives, are printed.

Interactively,          use           `ACECosetsThatStabiliseSubgroup'
(see~"ACECosetsThatStabiliseSubgroup").

\>`statistics'{option statistics}@{option `statistics'}
\>`stats'{option stats}@{option `stats'}&
Dump enumeration statistics.
(Shortest abbreviation of `statistics' is `stat'.)

If the statistics package is compiled into the {\ACE} code,  which  it
is by default (see the `options'~"option options" option),  then  this
option  dumps  the  statistics  accumulated  during  the  most  recent
enumeration. See the `enum.c' source  file  for  the  meaning  of  the
variables.

Interactively, use `ACEDumpStatistics' (see~"ACEDumpStatistics").

\>`style'{option style}@{option `style'}&
Print the current enumeration style.

This option prints the current enumeration style, as deduced from  the
current `ct' and `rt' parameters (see~"Enumeration Style").

Interactively, use `ACEStyle' (see~"ACEStyle").

\>`tw:=[<val>, <word>]'{option tw}@{option `tw'}
\>`trace:=[<val>, <word>]'{option trace}@{option `trace'}&
Traces `<word>' through the coset table, starting  at  coset  `<val>';
<val> must be a positive integer, and <word> must be  a  word  in  the
group generators.

This option prints the final coset number of the trace, if  the  trace
completes.

Interactively, use `ACETraceWord' (see~"ACETraceWord").

\enditems

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Options that Modify the Coset Table}

\beginitems

\>`recover'{option recover}@{option `recover'}
\>`contiguous'{option contiguous}@{option `contiguous'}&
`Recover' space used by dead coset numbers\index{dead coset (number)}.
(Shortest  abbreviation  of  `recover'   is   `reco',   and   shortest
abbreviation of `contiguous' is `contig'.)

This option invokes the compaction routine on the table to recover the
space used by any dead coset numbers. A `CO' message line  is  printed
if any cosets were recovered, and a  `co'  line  if  none  were.  This
routine is called automatically if  the  `cycles',  `nc',  `print'  or
`standard' options (see~"option cycles", "option nc",  "option  print"
and "option standard") are invoked.

Interactively, use `ACERecover' (see~"ACERecover").

\>`standard'{option standard}@{option `standard'}&
Compacts {\ACE}'s  coset  table  and  standardises  the  numbering  of
cosets. (Shortest abbreviation: `st'.)

For a given ordering of the generators in the columns of the table, it
produces a canonical numbering of the cosets. This function  does  not
display the new table; use the `print' (see~"option print") for  that.
Such a table has the property that a row-major scan  (i.e.~a  scan  of
the successive rows of the *body* of the table row by row,  from  left
to right) encounters previously unseen cosets in numeric order.

*Notes:*
In a canonic  table,  the  coset  representatives  are  ordered  first
according to length and then the lexicographic order  defined  by  the
order the generators and their inverses head the columns.  Note  that,
since {\ACE} avoids having an involutory generator in the first column
when it can, this lexicographic order does not necessarily  correspond
with the order in which the generators were first put to  {\ACE}.  Two
tables are equivalent only if their canonic forms are the same.  Also,
standardising the coset numbering within {\ACE} does *not* affect  the
{\GAP} coset table obtained via `ACECosetTable'.

*Guru Notes:*
In  half  of  the  ten  standard  enumeration   strategies   of   Sims
\cite{Sim94}, the table is standardised repeatedly. This is  expensive
computationally, but can result in fewer cosets being  necessary.  The
effect of doing this can be investigated  in  {\ACE}  by  (repeatedly)
halting  the  enumeration  (by  say,  imposing  timing  restrictions),
standardising the coset numbering, and continuing.

Interactively,             use             `ACEStandardCosetNumbering'
(see~"ACEStandardCosetNumbering").

\enditems

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Options for Comments}

\beginitems

\>`text:=<string>'{option text}@{option `text'}&
Prints <string> in the output; <string> must be a string.

This allows the user to add comments to the output from {\ACE}.

\>`aceincomment:=<string>'{option aceincomment}@{option `aceincomment'}&
Print comment <string> in the {\ACE} input; <string> must be a string.
(Shortest abbreviation: `aceinc'.)

This option prints the comment <string> behind a sharp sign (`\#')  in
the input to {\ACE}. Only useful  for  adding  comments  (that  {\ACE}
ignores) to standalone input files.

\enditems

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%E
